Interactive Summarization of Relational Tables

### 2014-10-29 TableSummarization/Paper/TableSummarization.tex
Added an experiment, and generated graph

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}

@@

\title{Interactive Summarization of Relational Tables}
% Can say multi-dimensional instead of relational
% Can say Dynamic Real-Time instead of Interactive?

@@

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints. The summary consists of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of that subset. We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table. Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
% TODO: Improve abstract.
\end{abstract}

@@

In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints. A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}. Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a rule-scoring black box as user input. In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.

\begin{example}\label{ex:introexample}
% Need justification for how investor got data. Or change example.
Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `size' denoting the number of non-$\star$ values of the rule for each rule.

%TODO: Change indentation format. All indentation at start, number of indents equal to depth of rule
\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}
% TODO: Should we display weight rather than size in table? After all, size is obvious. In intro example, we say that for this table, weight equals number of non-$\star$ columns.

@@

To summarize, we attempt to summarize relational data in a way that is:
\squishlist
\item \textbf{Flexible : } We allow users to specify a weighting function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
\item \textbf{Online : } Does not require any heavy pre-computation. 
\squishend

There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}. These papers mostly focuses on giving the most 'surprising' information to the user i.e. information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution. For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user. In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
% TODO: Say it's complementary. Also give an example where your thing is good. Mention MDL here. Or that max-entropy is hard to interpret for human, even though formally very clear. Our thing also works well when user doesn't know the set of column values. The set, at least number of column values, are needed for imagining the max-entropy distribution.

@@

\section{Preliminaries} 
\label{sec:preliminaries}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table. Our objective (formally defined later) is to summarise the table with a list of {\em rules}.

A {\em rule} is a tuple with one field per column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each field of the rule can either be one of the values in the corresponding column of the table, or $\star$, which is a wildcard character representing all values in the column. For a table column with numerical values, we can allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all fields. A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user.

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

@@

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule. {\em Count}($r$) is defined as the total number of tuples $t \in T$ that are covered by $r$. Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of MCount indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list. Let {\em Size}($r$) refer to the number of non-starred values in rule $r$. We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} \text{MCount}(r)W(r)$$.

@@

Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r)\text{MCount}(r)$$
subject to the following constraints:
\begin{enumerate}
\item If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{enumerate}

@@

\section{Weighting Rules}
\label{sec:weighting}
We want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is. Given this weighting, our total score for our rule-list is given by : $$\sum_{r\in R} \text{MCount}(r)W(r)$$. Let us consider some simple kinds of $W(r)$s and their meanings:
\begin{enumerate}
\item $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :

Here we set weight equal to the size of the rule. Consider the examples in Table \ref{table:sizescoringexample}. Weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus when we are going to marginalize on one extra attribute ($B$ in this case) it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an interpretation for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with MCount $100$, we are going to get a $100$ of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those. Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-set. Since having more of the table pre-filled is better, maximising score gives us a desirable set of rules.

\item $W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the 'complexity' of the column. The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes log($|c|$) bits of information. A non-starred valued in a rule $r$ thus pre-fills $\text{MCount}(r) \text{log}(|c|)$ bits of the table. Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.

Note that this scoring function is closely related to the Minimum Description Length (MDL) of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score. 
\end{enumerate}

@@

When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}. At any point, the user can click on either a rule, or a star, to 'expand' the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of sub-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule would display table~\ref{table:introexample}. Clicking further on the second rule would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows sub-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Table with only the trivial rule \label{table:trivial}}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. We now formally state the problem:

Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ sub-rules of $r^{\prime}$ that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)\text{MCount}(r)$$

@@

@@

Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. We want to find the $S$ that maximises $\text{Score}(S)$. We solve this part of the problem next.

\subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
We will now show that the Score function over sets of rules has a property called submodularity, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

@@

\title{Interactive Summarization of Relational Tables}

@@

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the most `interesting' summary of the table, subject to user-specified constraints. The summary consists of a list of `rules', where a rule is a pattern specifying values occurring in some columns of the table, along with the number of tuples that are covered by the rule. For instance, the rule $(a, b, \star, 1000)$ tell us that there are a $1000$ tuples with value $a$ in the first column and $b$ in the second column. We allow the user to specify what kind of rules to prefer in the summary via a black box weighting function, and to further selectively explore parts of the table. We describe an algorithm for finding the approximately optimal list of rules according to the user-specified weighting function, and a sampling based scheme for dealing with large tables. Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
\end{abstract}

@@

In this paper, we consider the following problem: Given a relational table, find the best list of `rules' to summarize the table, subject to user specified constraints. A `rule' is a pattern that describes a set of tuples in a relational table (defined more precisely in Section~\ref{sec:preliminaries}. Our summarization algorithm is highly {\em flexible}, allowing the user to specify what kind of rules are preferred, by taking a black box `weight function' as user input. In addition, the summary is {\em fast}, with a response time measured in seconds on our experimental datasets, and {\em interactive}, with a user interface that allows the user to explore parts of the table in greater detail.

\begin{example}\label{ex:introexample}
% Need justification for how investor got data. Or change example.
Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `weight' in this case denoting the number of non-$\star$ values of the rule for each rule. The weight function can in general be customized by the user.

%TODO: Change indentation format. All indentation at start, number of indents equal to depth of rule
\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}

@@

To summarize, we attempt to summarize relational data in a way that is:
\squishlist
\item \textbf{Flexible : } We allow users to specify a weight function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
\item \textbf{Online : } Does not require any heavy pre-computation. 
\squishend

There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}. These papers mostly focuses on giving the most 'surprising' information to the user i.e. information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution. For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user. In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible. Our approaches are thus complementary and serve different purposes.
% TODO: max-entropy is hard to interpret for human, even though formally very clear. Our thing also works well when user doesn't know the set of column values. The set, at least number of column values, are needed for imagining the max-entropy distribution.

@@

\section{Preliminaries} 
\label{sec:preliminaries}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table.

Our objective (formally defined later) is to summarize the table with a list of {\em rules}. A {\em rule} is a tuple with a value for each column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all columns.

A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values for all columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. 
%TODO: Give name to out system (and paper)

We say that rule $r_1$ {\em subsumes} rule $r_2$ if and only if $r_1$ has no more stars than $r_2$ and their values match wherever they both have non-starred values. For example, rule ($a$, $\star$) subsumes rule ($a$, $b$). Thus if $r_1$ subsumes $r_2$, then for all tuples $t$, $t \in r_2 \Rightarrow t \in r_1$. We also write $r_1 \geq r_2$ to denote that $r_1$ subsumes $r_2$, and $r_1 > r_2$ if $r_1$ subsumes but is not equal to $r_2$ i.e. there is at least one column $c$ such that $r_2$ has a non-$\star$ value in $c$ where $r_1$ does not. If $r_1$ subsumes $r_2$, we also say that $r_1$ is a {\em sub-rule} of $r_2$ and $r_2$ is a {\em super-rule} or $r_1$.

@@

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule.

We now define some additional properties of rules. Let $r$ be a rule. Then:
\squishlist
\item {\em Count}($r$) is the total number of tuples $t \in T$ that are covered by $r$. 
\item {\em MCount}($r$) (which stands for `Marginal Count') is the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user, to increase the coverage of the rule-list. 
\item {\em Size}($r$) refers to the number of non-starred values in rule $r$. 
\item We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. 
\squishend

The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} MCount(r)W(r)$$. Our problem is to choose the rule-list that maximises total score. We use $MCount$ rather than Count to increase coverage of the table. If we had defined total score as $$\sum_{r \in R} \text{Count}(r)W(r)$$, then our optimal rule-list would contain 
%TODO: Finish writing this part.

@@

\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r)MCount(r)$$
for one of the following interactions:
\begin{enumerate}
\item $[$Rule Interaction$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star Interaction$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{enumerate}
\end{problem}

@@

\section{Weighting Rules}
\label{sec:weighting}
We now describe how we score rules. At a high level, we want our rules to be as descriptive of the table as possible, i.e. given the rules, it should be as easy as possible to reproduce the table. We consider a general family of weighting functions, that assigns for each rule $r$, a weight $W(r)$ that depends on how expressive the rule is. Given this weight per rule, our total score for our rule-list is given by : $$\sum_{r\in R} MCount(r)W(r)$$. Let us consider some forms for the function $W(r)$ and their interpretation:
\begin{enumerate}
\item $W(r) = |\left\lbrace c \in C \mid r(c) \neq \star \right\rbrace |$ :

Here we set weight equal to the size of the rule. Consider the examples in Table \ref{table:sizescoringexample}. The weight for rule ($a$, $b_1$) is $2$, while weight for ($a$, $\star$) is $1$. Thus the total score for the rule-list with these two rules would be $2 \times 100 + 1 \times 900 = 1100$. If we replaced the rule ($a$, $\star$) by two rules ($a$, $b_2$) and ($a$, $\star$), then the $MCount$ of rule ($a$, $\star$) would reduce to $600$, since $300$ of its tuples would be covered by the previous rule $(a, b_2)$. Thus our total score would be $2 \times 100 + 2 \times 300 + 1 \times 600 = 1400$. If we had instead replaced ($a$, $\star$) by ($a$, $b_3$) and ($a$, $\star$), then our score would have been $1500$ which is $> 1400$. Thus when we marginalize on one extra column ($B$ in this case) and include a rule where that column is instantiated, it is better to do so on the value which occurs more frequently (in this case, $b_3$ which occurred $400$ times, compared to the $300$ of $b_2$).

To get an intuitive feel for this scoring function, imagine we are trying to reconstruct the table from the rules. Since we have rule ($a$, $b_1$) with $MCount$ $100$, we are going to get a $100$ of the table's tuples from this rule. For those hundred tuples, out of the $200$ total values to be filled ($2$ per tuple, since there are $2$ columns), all $200$ values will already have been filled (since the rule specifies both columns). Thus this rule contributes $200$ to the score. For the rule ($a$, $\star$), there are $900$ table tuples, and the $a$ value will be pre-filled for those. Thus $900$ slots of these tuples have been pre-filled, and so the rule contributes $900$ to the total. Thus this scoring function can be thought of as the number of values that have been pre-filled in the table by our rule-list. Since having more of the table pre-filled is better, maximizing the score gives us a desirable set of rules.

\item $W(r) = \sum_{c \in C : r(c) \neq \star} \text{log}_2(|c|)$ where $|c|$ refers to the number of distinct possible values in column $c$. Like the previous scoring function, this one adds some weight for every non-starred value in a rule. But instead of adding a weight of $1$ for every non-starred value, it varies the weight added depending on the `complexity' of the column. The reason behind this is the following: Say column $c_1$ is a boolean, while $c_2$ is a column with $20$ possible values. Then, a rule that gives us a value for $c_2$ is clearly giving us more information than a rule that gives us a value for $c_1$. Thus, this scoring function gives a higher weight to a rule that gives a value for a column with more distinct values.

The interpretation for this function is similar to the one for the last function. Once again, we can imagine trying to reconstruct the table from the rules, and look at how much of the table is pre-filled by the rules. But this time, we count the number of `bits' that are pre-filled. For a column $c$, specifying a value in the column takes log($|c|$) bits of information. A non-starred valued in a rule $r$ thus pre-fills $MCount(r) \text{log}(|c|)$ bits of the table. Hence this scoring function gives us the number of pre-filled bits of the table by the rule-set.

Note that this scoring function is closely related to the Minimum Description Length (MDL)~\cite{Grunwald:2007:MDL:1213810} of a table. If we describe a table using the rule-set, plus values to fill in for $\star$s in the rules to get tuples, then finding a set of rules that tries to minimize the length of this description, is equivalent to finding the rule-set that maximises the total score. 
\end{enumerate}

@@

When the user starts using our system, he sees a table with a single trivial rule, as shown in table~\ref{table:trivial}. At any point, the user can click on either a rule, or a star within a rule, to `expand' the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:trivial} would display table~\ref{table:introexample}. Clicking further on the second rule would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

Instead of clicking on a rule, the user can click on one of the $\star$s, say in column $c$ of rule $r$. This will also cause the rule $r$ to expand into a rule-list, but this time the new displayed rules are guaranteed to have non-$\star$ values for in column $c$. For instance, if the user clicks on the $\star$ in the product column of the walmart rule, he will see table~\ref{table:introexample3}, which shows super-rules of the walmart rule all specific to some product. This operation is useful if the user is more interested in a particular column that is unlikely to be instantiated in the top rules otherwise.

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
\end{tabular}
\caption{Table with only the trivial rule \label{table:trivial}}
\end{table}

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \cline{1-5}
$\triangleright$ target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & California & $80$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ walmart & bicycles & $\star$ & $150$ & $2$ \\  \hline
\end{tabular}
\caption{Result of clicking on a $\star$ \label{table:introexample3}}
\end{table}

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. If the user had clicked on a $\star$ in a column $c$, then we have the additional restriction that all resulting rules must have a non-$\star$ value in column $c$. We can reduce Problem~\ref{prob:optimal-rule-list} to a simpler problem given below:

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r)$$
\end{problem}

@@

Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\squishlist
\item $[$Rule Interaction$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we make on pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Interaction$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\squishend

@@

Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. This gives us 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r)$$
\end{problem}

The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.

\subsection{Greedy approximation algorithm for Problem~\ref{prob:optimal-rule-set}}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.
